A Minimal Demo of \texttt{glyphmaps}

### 2012-06-04 paper/paper.tex
tweaked geom_coxcomb. Wrote paper

\maketitle
This document demonstrates the basic features of the \texttt{ggplyr} package (name may change). \texttt{ggplyr} was written to create plots like the ones below. Each plot was made with \texttt{ggplyr}. The first two plots recreate plots from \citet{wickham2011split}. The last two recreate plots from \citet{wickham2012glyph}.

\section{Glyphs}
The common feature of these graphs is that they are built around glyphs. Glyphs are geometric objects (i.e, geoms) designed to display information within each geom. In other words, a glyph can display information even if it ids drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

Glyphs reveal a heirarchical structure to graphics: every plot is a collection of geoms, each of which can be thought of as its own self contained plot. Sometimes these subplots are not very interesting, as in the subplot created by a single point geom. Sometimes, they are quite complex, as in the star glyph of Figure~\ref{fig:star-glyphs}.

Graphs inherit this heirarchical structure from the data they describe. Data is produced through an iterative process of collecting observations, grouping observations and summarizing groups of observations to create more compact, information dense sets of data. Humans innately perform this process when collecting data.  It is a cognitive pattern of the human brain which I will write about in the second cognitive chapter of my thesis.

Glyph maps simultaneously expose data from multiple levels of this process. As a set of geoms, the glyphs reveal relationships between data points in the higher level, compact data set. As individual plots, each glyph retains information about the data points in the lower level group of data that it summarises. This dual display makes glyph maps particularly useful for certain data analysis tasks. It also provides two different approaches to constructing glyph maps.

Glyph maps can be built from the top down by treating each glyph as a geom within the plot of interest. \texttt{ggplyr} provides new geoms \texttt{geom_star}, \texttt{geom_radar}, \texttt{geom_dart}, and \texttt{geom_plyr} which allow a user to quickly incorporate geoms into an existing plot.

Alternatively, glyph maps can be built from the bottom up by treating each glyph as an existing plot and then combining these plots together with an external coordinate system. This approach parallels the split-apply-combine strategy of data analysis \citep{wickham2011split}. \texttt{ggplyr} provides a set of functions such as \texttt{ggply}, \texttt{dgply}, \texttt{grid}, and \texttt{nest} that help users use the familiar split-apply-combine process to create complex glyph maps. This approach will illustrate that any graph can be turned into a glyph to be used in a higher order graph.

The remainder of this paper will demonstrate how users can use \texttt{ggplyr} tools to build glyph maps from either the bottom up or the top down.

Although we have chosen to describe such plots as glyph maps, they really represent a general class of graphs that have been difficult to construct until now. Because each glyph is equivalent to an independent plot, glyph maps describe any plot that embeds other plots within itself. Embedded plots occur in a variety of contexts and I have a stack of photocopies to help demonstrate this. Embedded plots also fit into the grammar of graphics, where they represent an extension of facetting. When plots are embedded into two categorical axes, a facetted plot results. When a graph is facetted on continuous axes (which is rarely done), a glyph map results.

(Interesetingly most embedded plots have settled to or are stuck in the stage of embedding plots within a geographic map. This resembles how Wainer, Tufte and others have traced the development of graphs through a concrete stage where cartesian axes seldom refered to anything other than latitude and longitude. Eventually graphs became more sophisticated and axes were used to represent  abstract concepts. Infovis developed from SciVis in the same way. Scivis focused on visualizing physical dimensions and eventually spawned InfoVis, which focuses on visualizing abstract dimensions. Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one of them to have a concrete interpretation.)

\section{top down methods}
\texttt{ggplyr}'s top down methods help users fit glyphs into existing methods of visualization. These methods provide new geoms (often based on new grobs) that can be used alongside existing geoms in \texttt{ggplot2}.

Top down methods are difficult to implement in \texttt{ggplot2} for two reasons
\begin{ennumerate}
  \item First, \texttt{ggplot2} calculates aesthetics on the entire data set at once. However, glyphs contain aesthetics that must be keyed to subsets of the data.
  \item Second, the final width and height of individual glyphs often depends on non-position aesthetics, such as angle and length. In the \texttt{ggplot2} pipeline, these aesthetics are scaled right before the plot ranges are trained. As a result, the final widths of glyphs must be computed at draw time and frequently place parts of the glyph outside the plot window.
\end{ennumerate}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{set.seed}\hlkeyword{(}\hlnumber{1121}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlkeyword{(}\hlsymbol{x}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{rnorm}\hlkeyword{(}\hlnumber{20}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
##  [1]  0.14496  0.43832  0.15319  1.08494  1.99954 -0.81188  0.16027  0.58589  0.36009
## [10] -0.02531  0.15088  0.11008  1.35968 -0.32699 -0.71638  1.80977  0.50840 -0.52746
## [19]  0.13272 -0.15594
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## [1] 0.3217
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{var}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## [1] 0.5715
\end{verbatim}
\end{kframe}
\end{knitrout}

The first element of \texttt{x} is 0.145. Boring boxplots
and histograms recorded by the PDF device:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }two{\ }plots{\ }side{\ }by{\ }side{\ }(option{\ }fig.show=\usebox{\hlnormalsizeboxsinglequote}hold\usebox{\hlnormalsizeboxsinglequote})}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{par}\hlkeyword{(}\hlargument{mar}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{cex.lab}{\ }\hlargument{=}{\ }\hlnumber{0.95}\hlkeyword{,}{\ }\hlargument{cex.axis}{\ }\hlargument{=}{\ }\hlnumber{0.9}\hlkeyword{,}{\ }\hlargument{mgp}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{2}\hlkeyword{,}{\ }\hlnumber{0.7}\hlkeyword{,}{\ }\hlnumber{0}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{tcl}{\ }\hlargument{=}{\ }\hlkeyword{-}\hlnumber{0.3}\hlkeyword{,}{\ }\hlargument{las}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots1} \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots2}

@@

Do the above chunks work? You should be able to compile the \TeX{}
document and get a PDF file like this one: \url{https://github.com/downloads/yihui/knitr/knitr-minimal.pdf}.
The Rnw source of this document is at \url{https://github.com/yihui/knitr/blob/master/inst/examples/knitr-minimal.Rnw}.

>>>>

\maketitle
This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change). \texttt{glyphmaps} was written to create plots like the ones below, which recreate plots similar to those in \citet{wickham2011split} and \citet{wickham2012glyph}. Both plots were made with \texttt{glyphmaps}. To run the examples in this document please download the \texttt{glyphmaps} package with the code below.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{install\usebox{\hlnormalsizeboxunderscore}github}\hlkeyword{(}\hlstring{"{}ggplyr"{}}\hlkeyword{,}{\ }\hlstring{"{}garrettgman"{}}\hlkeyword{,}{\ }\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{glyphmaps}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}us}{\ }\hlkeyword{+}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{grid}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}line}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{time}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{pred}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{grid.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{lon}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{x.nbin}{\ }\hlargument{=}{\ }\hlnumber{58}\hlkeyword{,}{\ }\hlargument{y.nbin}{\ }\hlargument{=}{\ }\hlnumber{25}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{avg}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{alpha}{\ }\hlargument{=}{\ }\hlnumber{0.2}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlnumber{NA}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=1.05\textwidth]{figure/minimal-glyphmaps-fig-11}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}nasa}{\ }\hlkeyword{+}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}star}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{r}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{date}\hlkeyword{,}{\ }\hlargument{x}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{major.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}long"{}}\hlkeyword{,}{\ }\hlstring{"{}lat"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.8\textwidth]{figure/minimal-split-fig}

\texttt{glyphmaps} helps users create embedded graphics: plots that have subplots built into them. The package is built around three main functions: \texttt{grid}, \texttt{glyph}, and \texttt{ply$\_$aes}. This demo will demonstrate these three functions, their options, and the various uses of embedded graphics.

\section{Glyphs}
Glyphs are geometric objects (i.e, geoms) that display information within the geom. In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-star-glyphs1} \includegraphics[width=0.49\textwidth]{figure/minimal-star-glyphs2}

\caption['Individual star glyphs comprise a complete polar plot with an internal coordinate system (r = temperature and angle = date)]{'Individual star glyphs comprise a complete polar plot with an internal coordinate system (r = temperature and angle = date). Multiple star glyphs can be organized with an external coordinate sytem to reveal second order effects.'\label{fig:star-glyphs}}
\end{figure}

Glyphs provide one connection between the grammar and graphics and embedded plots. Every glyph is an embedded plot and each embedded plot is a type of glyph. Hence, we can treat embedded plots as a type of geom. Embedded plots can also be thought of as facetting on a continuous scale. However, since embedded plots are often used with other non-facetted layers, it is easier to implement embedded plots as a type of geom.

\subsection{glyph}
\texttt{glyph} provides a general method for creating glyph geoms in \texttt{ggplot2}. \texttt{glyph} works in conjunction with \texttt{ggplot2} layer objects in the following way. Each layer specifies a type of graph through a combination of geom, stat, aesthetics, parameters, and a position adjustment. \texttt{glyph} uses this specification to build the individual glyphs (i.e, \texttt{glyph} builds glyphs that match the plot specified by the layer). \texttt{glyph} will plot a separate glyph for different subsets of the layer's data. Users specify how to split the data into subsets with the \texttt{glyph.by} argument. This argument behaves similarly to the \texttt{.variables} argument in the \texttt{plyr} package \citep{wickham2011split}. Finally, \texttt{glyph} plots each glyph on the x and y axis specified by the \texttt{major.aes} argument. These ``major axiis'' are mapped to the data just as other \texttt{ggplot2} aesthetics are mapped to the data. Hence they should be specified with the \texttt{aes} function.

To summarize, a layer can be turned into a set of glyphs with the glyph function and three arguments:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlsymbol{points}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{surftemp}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{2}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }A{\ }normal{\ }plot}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{points}\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }Glyphed{\ }plot}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlargument{layer}{\ }\hlargument{=}{\ }\hlsymbol{points}\hlkeyword{,}{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-glyph-example1} \includegraphics[width=0.49\textwidth]{figure/minimal-glyph-example2}

\texttt{glyph} gives users incredible freedom to design their own types of glyphs. Any graph that can be specified with a single layer in \texttt{ggplot2} can be turned into a glyph. \texttt{glyphmaps} also provides two new geoms that can be used to create frequently seen glyphs. These are \texttt{geom$\_$star} and \texttt{geom$\_$coxcomb}, see Figure~\ref{fig:star}. Although coxcomb glyphs don't seem much better than Venn pie-agrams (which I suppose we could also do with glyphmaps).
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}star}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{r}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{date}\hlkeyword{,}{\ }\hlargument{x}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{ozone}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{alpha}{\ }\hlargument{=}{\ }\hlnumber{0.5}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}coxcomb}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{state}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-star1} \includegraphics[width=0.49\textwidth]{figure/minimal-star2}

\texttt{glyph} also has a variety of optional arguments that allow users to customize glyph plots. These allow the user to control the dimensions of each glyph, to control the scale within each glyph, to add reference objects to each glyph, and to merge glyphs that overlap into a single glyph.

\subsubsection{width, height}
The \texttt{width} and \texttt{height} arguments of \texttt{glyph} control the width and height of each glyph in the plot.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-dims1} \includegraphics[width=0.49\textwidth]{figure/minimal-dims2}

\subsubsection{x$\_$scale, y$\_$scale}
The \texttt{x$\_$scale} and \texttt{y$\_$scale} arguments of \texttt{glyph} control the scaling within each glyph. As with facets, glyphs share the same scaling by default. The x and y scales can also be set to `free'.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }glyphs{\ }share{\ }y{\ }scale}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }independent{\ }y{\ }scales}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{y\usebox{\hlnormalsizeboxunderscore}scale}{\ }\hlargument{=}{\ }\hlsymbol{free}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-scales1} \includegraphics[width=0.49\textwidth]{figure/minimal-scales2}

\subsubsection{reference objects}
Users can set the \texttt{reference} argument to include reference objects with each glyph. These objects provide a frame of reference which facilitates comparing between glyphs. The \texttt{reference} argument should be set to one of \texttt{ref$\_$box}, \texttt{ref$\_$hline}, \texttt{ref$\_$vline}, \texttt{ref$\_$points}, or \texttt{NULL}. Each of these functions creates a different type of reference object, see Figure~\ref{fig:refs}. Each also accepts a mapping and parameters to customize the appearance of the reference objects.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }boxes}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}\hspace*{\fill}\\
\hlstd{}{\ }{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hspace*{\fill}\\
\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }hlines}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
